---
output: github_document
---

# ratSpermBSseqData https://img.shields.io/badge/data-package-blue.svg

## Overview

`ratSpermBSseqData` is an **R data package** that provides the
[`BSseq`](https://bioconductor.org/packages/release/bioc/html/bsseq.html) object with the methylation data by [Altintas et al. 2021](https://faseb.onlinelibrary.wiley.com/doi/10.1096/fj.202100215R) from the GEO database
[GSE173898](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE173898). 

Placebo (Ctrl) and nicotine (Nico) treated F0 mothers. Sperm cells are isolated from F1 generation (n = 10 per group).

The dataset was constructed from Bismark coverage files using `bsseq::read.bismark()`.
---

## Installation

You can install the development version from GitHub using one of the following methods:

```{r install_pak}
if (!require("pak")) install.packages("pak")
#pak::pak("altintasali/ratSpermBSseqData")
```

# Usage 

## Loading data

```{r load_data} 
library(ratSpermBSseqData)

# Load the BSseq object
data("bsseq_rat_sperm")

# Inspect the object
bsseq_rat_sperm
```

## Extract information

```{r extract_data}

library(bsseq)
cov <- getCoverage(bsseq_rat_sperm, type = "Cov")
meth <- getCoverage(bsseq_rat_sperm, type = "M")

dim(cov)    # rows = CpG loci, columns = samples
head(cov)
```
