---
output: github_document
---
<!-- badges: start -->
![Data Package](https://img.shields.io/badge/data-package-blue)
<!-- badges: end -->

# ratSpermBSseqData 

## Overview

`ratSpermBSseqData` is an **R data package** that provides the
[`BSseq`](https://bioconductor.org/packages/release/bioc/html/bsseq.html) object with the methylation data by [Altintas et al. 2021](https://faseb.onlinelibrary.wiley.com/doi/10.1096/fj.202100215R) from the GEO database
[GSE173898](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE173898). 

Placebo (Ctrl) and nicotine (Nico) treated F0 mothers. Sperm cells are isolated from F1 generation (n = 10 per group).

The dataset was constructed from Bismark coverage files using `bsseq::read.bismark()`.

## Installation

You can install the development version from GitHub using one of the following methods:

```{r install_pakwarning=FALSE, message=FALSE}
if(!require("remotes")){install.packages("remotes")}
if(!require("ratSpermBSseqData")){
  remotes::install_github("altintasali/ratSpermBSseqData", 
                          dependencies = TRUE)
}
```

# Usage 

## Loading data

```{r load_data, warning=FALSE, message=FALSE} 
library(ratSpermBSseqData)

# Load the BSseq object
data("bsseq_rat_sperm")

# Inspect the object
bsseq_rat_sperm
```

## Extract information

```{r extract_data, warning=FALSE, message=FALSE}
library(bsseq)
cov <- getCoverage(bsseq_rat_sperm, type = "Cov")
meth <- getCoverage(bsseq_rat_sperm, type = "M")

dim(cov)    # rows = CpG loci, columns = samples
head(cov)
```
